---
title: "Computation of Greeks under rough Volterra stochastic volatility models using the Malliavin calculus approach"
author:
  - Neil Mascarenhas
  - Zijie Gu
date: today
institute: "Cornell University"
format:
  beamer:
    toc: true
    toc-depth: 1
    toc-title: "Contents"
    theme: Madrid
    colortheme: seahorse
    aspectratio: 169
    slide-level: 3
    include-in-header:
      text: |
        \useoutertheme[compress]{miniframes}
        \setbeamertemplate{headline}{%
          \begin{beamercolorbox}[ht=2.2ex,dp=0.8ex,center]{section in head/foot}
            \insertsectionnavigationhorizontal{\paperwidth}{}{}
          \end{beamercolorbox}%
          \begin{beamercolorbox}[ht=2.2ex,dp=0.8ex,center]{subsection in head/foot}
            \insertsubsectionnavigationhorizontal{\paperwidth}{}{}
          \end{beamercolorbox}%
        }
        \setbeamercolor{mini frame}{fg=structure.fg,bg=structure.fg}
        \setbeamercolor{mini frame in current subsection}{fg=structure.fg,bg=structure.fg}
        \setbeamercolor{mini frame in other subsection}{fg=structure.fg,bg=white}
        \AtBeginSubsection[]{}
        \setbeamertemplate{section page}{%
          \begin{centering}
            \begin{beamercolorbox}[sep=12pt,center,rounded=true,shadow=true]{part title}
              \usebeamerfont{section title}\insertsection\par
            \end{beamercolorbox}
          \end{centering}
        }
---

# Background

## Background

### Background

For a discounted payoff

$$
V^\theta = \mathbb{E}\left[e^{-rT} f(S_T^\theta)\right],
$$

a Greek is the partial derivative:

$$
\partial_\theta V^\theta
= \partial_\theta \mathbb{E}[e^{-rT} f(S_T^\theta)]
$$

The project compares three estimator families:

- Finite difference: generic, but sensitive to bump size.
- Pathwise derivative: efficient, but needs enough payoff smoothness.
- Malliavin / likelihood-ratio style: moves the derivative from the payoff to a random weight.

## Finite Difference

### Finite Difference

$\partial_\theta V^\theta$ can be approximated by
$$
\partial_\theta V^\theta\approx\frac{V^{\theta+h}-V^{\theta-h}}{2h}
$$

This discretization has bias $O(h^2)$ and variance $O(h^{-2})$, and therefore
- too large $h$ creates discretization bias;
- too small $h$ amplifies Monte Carlo noise.

The second order derivative is approximated by
$$
\partial^2_{\theta\theta} V^\theta\approx\frac{V^{\theta+h}-2V^\theta+V^{\theta-h}}{h^2}
$$

This discretization has bias $O(h^2)$ and variance $O(h^{-4})$ and can be very unstable.

## Pathwise Derivative

### Pathwise Derivative

If the payoff function $f$ is *nice* enough, we can swap the expectation and derivative
$$
\partial_\theta \mathbb{E}[e^{-rT} f(S_T^\theta)]=\mathbb{E}[\partial_\theta e^{-rT} f(S_T^\theta)]
$$

> $f$ is not *nice* if $f$ has discontinuity.

The pathwise method is very efficient and accurate (0 bias), when applicable.

## Malliavin Approach

### Likelihood Ratio Method and the Malliavin Approach

If the terminal state has density $p_\theta(x)$, then we have
$$
\partial_\theta \mathbb{E}[f(X^\theta)]=\mathbb{E}[f(X^\theta)\frac{\partial_\theta p_\theta(X^\theta)}{p_\theta(X^\theta)}]=\mathbb{E}_\theta[f(X^\theta)\partial_\theta \log p_\theta(X^\theta)]
$$

This is essentially computing a weighted ($\log p_\theta(X^\theta)$) mean of all the simulated payoffs.

It is also very efficient and accurate (0 bias), provided we know $p_\theta(x)$ explicitly.

Malliavin's method extends this method via the integration-by-parts formula:
$$
\partial_\theta \mathbb{E}[f(X^\theta)]=\mathbb{E}\Bigg[f(X^\theta)\cdot\delta\Bigg(\frac{\partial_\theta X^\theta h}{\int_0^Th_t\mathcal{D}_tX^\theta\mathrm{d}t}\Bigg)\Bigg]
$$

# Numerical Experiment

## Models

### Numerical Experiment

We test Malliavin estimators for $\Delta$ and $\Gamma$ under two rough Volterra volatility models:

\begin{columns}[T,totalwidth=\textwidth]
\begin{column}{0.48\textwidth}
\textbf{$\alpha$-RFSV / rough Bergomi}
$$
V_t = V_0\exp\left\{\xi Y_t-\frac12\alpha\xi^2 r(t)\right\},
\qquad \alpha=1.
$$
\end{column}
\begin{column}{0.48\textwidth}
\textbf{Rough Stein--Stein}
$$
dV_t=\kappa(\theta-V_t)\,dt+\nu\,dY_t.
$$
\end{column}
\end{columns}

\vspace{0.4em}

In both cases,
$$
Y_t=\int_0^t K_H(t,s)\,dZ_s,\qquad
Z_s=\rho W_s+\sqrt{1-\rho^2}\widetilde W_s.
$$

## Estimator Comparison

### Estimator Comparison

\begin{itemize}
\item \textbf{Malliavin:} estimate Greeks as discounted payoffs multiplied by model-specific Malliavin weights.
\item \textbf{Finite difference:} central difference in $S_0$ using common random numbers.
\item \textbf{Pathwise:} included for vanilla $\Delta$, where payoff differentiability permits it.
\end{itemize}

\vspace{0.4em}

The convergence plots report running Monte Carlo means for $\Delta$ and $\Gamma$ with 99\% confidence intervals.

# Experiment Setup

## Overview

### Experiment Setup

\begin{itemize}
\item Simulate terminal prices under both rough volatility models using the same rough-driver structure.
\item Estimate $\Delta$ and $\Gamma$ for vanilla, digital, and gap payoffs.
\item Compare Malliavin weights with finite differences; add pathwise $\Delta$ where the payoff is smooth enough.
\end{itemize}

## Simulation

### Simulation

\begin{itemize}
\item Shared parameters: $S_0=100$, $r=0.05$, $T=1$, $H=0.15$, $\rho=-0.756$, $V_0=0.354$, $\varepsilon=10^{-6}$.
\item $\alpha$-RFSV parameters: $\alpha=1$, $\xi=0.212$.
\item Rough Stein--Stein parameters: $\kappa=2.0$, $\theta=0.5$, $\nu=0.3$.
\item Monte Carlo design: $200{,}000$ paths, $2{,}000$ time steps, fixed seed, and log-spaced checkpoints from $1{,}000$ paths onward.
\item Finite difference bump: $h=10^{-5}$ applied to $S_0$ with the same simulated volatility and Brownian paths.
\end{itemize}

## Payoffs

### Payoffs

\begin{columns}[T,totalwidth=\textwidth]
\begin{column}{0.48\textwidth}
\textbf{Continuous payoff}

\begin{itemize}
\item Vanilla call/put with $K=100$.
\item Compare Malliavin, finite difference, and pathwise $\Delta$.
\end{itemize}
\end{column}
\begin{column}{0.48\textwidth}
\textbf{Discontinuous payoff}

\begin{itemize}
\item Digital call/put with $K=100$.
\item Gap call with $(K_1,K_2)=(110,100)$.
\item Gap put with $(K_1,K_2)=(90,100)$.
\item Compare Malliavin against finite difference.
\end{itemize}
\end{column}
\end{columns}

\vspace{0.4em}

For rough Stein--Stein, paths with near-zero Malliavin denominator are filtered for numerical stability.

## Vanilla Payoffs

### Vanilla Options

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 4.4
#| out-width: "100%"

S <- seq(50, 150, length.out = 1000)
K <- 100

vanilla_call <- pmax(S - K, 0)
vanilla_put <- pmax(K - S, 0)

old_par <- par(mfrow = c(1, 2), mar = c(4, 4, 2, 1), oma = c(0, 0, 0, 0))

plot(S, vanilla_call, type = "l", lwd = 3, col = "#1f77b4",
     xlab = expression(S[T]), ylab = "Payoff",
     main = "Vanilla Call")
abline(v = K, lty = 2, col = "gray40")
grid()

plot(S, vanilla_put, type = "l", lwd = 3, col = "#d62728",
     xlab = expression(S[T]), ylab = "Payoff",
     main = "Vanilla Put")
abline(v = K, lty = 2, col = "gray40")
grid()

par(old_par)
```

## Digital Payoffs

### Digital Options

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 4.4
#| out-width: "100%"

S <- seq(50, 150, length.out = 1000)
K <- 100

digital_call <- as.numeric(S > K)
digital_put <- as.numeric(S < K)

old_par <- par(mfrow = c(1, 2), mar = c(4, 4, 2, 1), oma = c(0, 0, 0, 0))

plot(S, digital_call, type = "n",
     xlab = expression(S[T]), ylab = "Payoff",
     main = "Digital Call", ylim = c(-0.05, 1.05))
grid()
lines(c(min(S), K), c(0, 0), lwd = 3, col = "#1f77b4")
lines(c(K, max(S)), c(1, 1), lwd = 3, col = "#1f77b4")
points(K, 0, pch = 1, cex = 1.6, lwd = 2, col = "#1f77b4")
points(K, 1, pch = 19, cex = 1.6, col = "#1f77b4")

plot(S, digital_put, type = "n",
     xlab = expression(S[T]), ylab = "Payoff",
     main = "Digital Put", ylim = c(-0.05, 1.05))
grid()
lines(c(min(S), K), c(1, 1), lwd = 3, col = "#d62728")
lines(c(K, max(S)), c(0, 0), lwd = 3, col = "#d62728")
points(K, 0, pch = 1, cex = 1.6, lwd = 2, col = "#d62728")
points(K, 1, pch = 19, cex = 1.6, col = "#d62728")

par(old_par)
```

## Gap Payoffs

### Gap Options

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 4.4
#| out-width: "100%"

S <- seq(50, 150, length.out = 1000)
K2 <- 100
K1_call <- 110
K1_put <- 90

gap_call <- as.numeric(S > K1_call) * pmax(S - K2, 0)
gap_put <- as.numeric(S < K1_put) * pmax(K2 - S, 0)

old_par <- par(mfrow = c(1, 2), mar = c(4, 4, 2, 1), oma = c(0, 0, 0, 0))

plot(S, gap_call, type = "n",
     xlab = expression(S[T]), ylab = "Payoff",
     main = "Gap Call")
abline(v = K2, lty = 3, col = "gray55")
grid()
lines(c(min(S), K1_call), c(0, 0), lwd = 3, col = "#1f77b4")
lines(S[S >= K1_call], S[S >= K1_call] - K2, lwd = 3, col = "#1f77b4")
points(K1_call, 0, pch = 1, cex = 1.6, lwd = 2, col = "#1f77b4")
points(K1_call, K1_call - K2, pch = 19, cex = 1.6, col = "#1f77b4")

plot(S, gap_put, type = "n",
     xlab = expression(S[T]), ylab = "Payoff",
     main = "Gap Put")
abline(v = K2, lty = 3, col = "gray55")
grid()
lines(S[S <= K1_put], K2 - S[S <= K1_put], lwd = 3, col = "#d62728")
lines(c(K1_put, max(S)), c(0, 0), lwd = 3, col = "#d62728")
points(K1_put, 0, pch = 1, cex = 1.6, lwd = 2, col = "#d62728")
points(K1_put, K2 - K1_put, pch = 19, cex = 1.6, col = "#d62728")

par(old_par)
```

## Model Paths

### Simulated Price and Volatility Paths

\begin{figure}
\centering
\begin{minipage}{0.495\textwidth}
\centering
\includegraphics[width=\linewidth,height=0.82\textheight,keepaspectratio]{www/stein_path.png}
\end{minipage}%
\begin{minipage}{0.495\textwidth}
\centering
\includegraphics[width=\linewidth,height=0.82\textheight,keepaspectratio]{www/rough_vol_path.png}
\end{minipage}
\end{figure}

# Results

## Vanilla Call

### Vanilla Call Options

\begin{figure}
\centering
\begin{minipage}{0.495\textwidth}
\centering
\includegraphics[width=\linewidth,height=0.82\textheight,keepaspectratio]{www/vanilla_call_stein.png}
\end{minipage}%
\begin{minipage}{0.495\textwidth}
\centering
\includegraphics[width=\linewidth,height=0.82\textheight,keepaspectratio]{www/vanilla_call_RFSV.png}
\end{minipage}
\end{figure}

## Vanilla Put

### Vanilla Put Options

\begin{figure}
\centering
\begin{minipage}{0.495\textwidth}
\centering
\includegraphics[width=\linewidth,height=0.82\textheight,keepaspectratio]{www/vanilla_put_stein.png}
\end{minipage}%
\begin{minipage}{0.495\textwidth}
\centering
\includegraphics[width=\linewidth,height=0.82\textheight,keepaspectratio]{www/vanilla_put_RFSV.png}
\end{minipage}
\end{figure}

## Digital Call

### Digital Call Options

\begin{figure}
\centering
\begin{minipage}{0.495\textwidth}
\centering
\includegraphics[width=\linewidth,height=0.82\textheight,keepaspectratio]{www/digital_call_stein.png}
\end{minipage}%
\begin{minipage}{0.495\textwidth}
\centering
\includegraphics[width=\linewidth,height=0.82\textheight,keepaspectratio]{www/digital_call_RFSV.png}
\end{minipage}
\end{figure}

## Digital Put

### Digital Put Options

\begin{figure}
\centering
\begin{minipage}{0.495\textwidth}
\centering
\includegraphics[width=\linewidth,height=0.82\textheight,keepaspectratio]{www/digital_put_stein.png}
\end{minipage}%
\begin{minipage}{0.495\textwidth}
\centering
\includegraphics[width=\linewidth,height=0.82\textheight,keepaspectratio]{www/digital_put_RFSV.png}
\end{minipage}
\end{figure}

## Gap Call

### Gap Call Options

\begin{figure}
\centering
\begin{minipage}{0.495\textwidth}
\centering
\includegraphics[width=\linewidth,height=0.82\textheight,keepaspectratio]{www/gap_call_stein.png}
\end{minipage}%
\begin{minipage}{0.495\textwidth}
\centering
\includegraphics[width=\linewidth,height=0.82\textheight,keepaspectratio]{www/gap_call_RFSV.png}
\end{minipage}
\end{figure}

## Gap Put

### Gap Put Options

\begin{figure}
\centering
\begin{minipage}{0.495\textwidth}
\centering
\includegraphics[width=\linewidth,height=0.82\textheight,keepaspectratio]{www/gap_put_stein.png}
\end{minipage}%
\begin{minipage}{0.495\textwidth}
\centering
\includegraphics[width=\linewidth,height=0.82\textheight,keepaspectratio]{www/gap_put_RFSV.png}
\end{minipage}
\end{figure}

# Conclusion

## Summary

### Conclusion

\begin{itemize}
\item Malliavin provides unbiased estimates for second order Greeks where traditional methods fail.
\item Finite difference becomes high variance and unstable for second order Greeks on exotic options with rough volatility.
\item Finite difference also provides poor estimates for first order Greeks on exotic options with rough volatility.
\item Pathwise derivative is efficient for smooth vanilla deltas, but fails for discontinuous payoffs.
\end{itemize}

## References

### References

\footnotesize

\begin{itemize}
\item Al-Foraih, M., Bures, O., Pospisil, J., \& Vives, J. (2025). \textit{Computation of Greeks under rough Volterra stochastic volatility models using the Malliavin calculus approach}. arXiv:2312.00405v2.
\end{itemize}
