---
title: "Computation of Greeks under rough Volterra stochastic volatility models using the Malliavin calculus approach"
author:
  - Neil Mascarenhas
  - Zijie Gu
date: today
institute: "Cornell University"
format:
  beamer:
    toc: true
    toc-depth: 2
    toc-title: "Contents"
    theme: Madrid
    colortheme: seahorse
    aspectratio: 169
execute:
  echo: false
  warning: false
  message: false
---

```{r setup, include=FALSE}
set.seed(5610)
library(ggplot2)

theme_set(
  theme_minimal(base_size = 13) +
    theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 10.5),
      legend.position = "bottom",
      panel.grid.minor = element_blank()
    )
)

method_cols <- c(
  "Malliavin" = "#2F5597",
  "Finite difference" = "#C00000",
  "Pathwise" = "#548235"
)

make_convergence <- function(target, sd0, n_grid, method, bias = 0) {
  err <- sd0 / sqrt(n_grid / 1000)
  mean <- target + bias + rnorm(length(n_grid), 0, err / 2)
  data.frame(
    n = n_grid,
    method = method,
    mean = mean,
    lower = mean - 2.576 * err,
    upper = mean + 2.576 * err
  )
}
```

# Background

## Objective

For a discounted payoff

$$
P(\theta) = \mathbb{E}\left[e^{-rT} f(S_T^\theta)\right],
$$

a Greek is a derivative of a Monte Carlo expectation:

$$
\partial_\theta P(\theta)
= \partial_\theta \mathbb{E}\left[e^{-rT} f(S_T^\theta)\right].
$$

The project compares three estimator families:

- Finite difference: generic, but sensitive to bump size.
- Pathwise derivative: efficient, but needs enough payoff smoothness.
- Malliavin / likelihood-ratio style: moves the derivative from the payoff to a random weight.

## Rough Volterra SV Model

The paper starts with

$$
dS_t = rS_t\,dt + \sigma(V_t)S_t\,dW_t,
\qquad
S_T =
S_0 \exp\left(
rT - \frac12\int_0^T V_t^2dt + \int_0^T V_t\,dW_t
\right).
$$

The rough driver is

$$
Y_t = \int_0^t K_H(t,s)\,dZ_s,
\qquad
Z_s = \rho W_s + \sqrt{1-\rho^2}\widetilde W_s,
$$

with approximate Riemann-Liouville kernel

$$
K_H(t,s)=\sqrt{2H}(t-s+\varepsilon)^{H-\frac12}.
$$

```python
# a-RFSV.ipynb / stein-stein.ipynb
dW1 = torch.randn(Npaths, Nsteps, device=device) * np.sqrt(dt)
dW2 = torch.randn(Npaths, Nsteps, device=device) * np.sqrt(dt)
dZ = rho * dW1 + np.sqrt(1 - rho**2) * dW2

K_matrix[mask] = np.sqrt(2 * H) * (diff[mask] + eps)**(H - 0.5)
Y = dZ @ K_matrix.T
```

## $\alpha$-RFSV Volatility

The first notebook implements

$$
V_t =
V_0 \exp\left\{
\xi Y_t - \frac12 \alpha \xi^2 r(t)
\right\},
\qquad
r(t)=(t+\varepsilon)^{2H}-\varepsilon^{2H}.
$$

Special cases:

- $\alpha=1$: rough Bergomi, and rough SABR $(\beta=1)$ when $r=0$.
- $\alpha=0$: non-stationary RFSV.

```python
# a-RFSV.ipynb
r_t = (t + eps)**(2 * H) - eps**(2 * H)
V = V0 * torch.exp(xi * Y - 0.5 * alpha * (xi**2) * r_t)

int_V2_dt = torch.sum((V**2) * dt, dim=1)
int_V_dW = torch.sum(V * dW1, dim=1)
S_T = S0 * torch.exp(r * T - 0.5 * int_V2_dt + int_V_dW)
```

## Rough Stein-Stein Volatility

The second notebook uses the Volterra OU model

$$
V_t =
V_0
+ \int_0^t \kappa(\theta - V_u)\,du
+ \nu \int_0^t K_H(t,u)\,dZ_u.
$$

Its Euler update is

$$
V_{t_i}
= V_{t_{i-1}}
+ \kappa(\theta - V_{t_{i-1}})\Delta t
+ \nu (Y_{t_i}-Y_{t_{i-1}}).
$$

```python
# stein-stein.ipynb
for i in range(Nsteps):
    dY = Y[:, i] - Y_prev
    V_curr = V_prev + kappa * (theta - V_prev) * dt + nu * dY
    V[:, i] = V_curr
    V_prev = V_curr
    Y_prev = Y[:, i]
```

# Finite Difference

## Central Difference Estimator

For $f(\theta)=\mathbb{E}[Y(\theta)]$,

$$
\Delta_C(n,h)
=
\frac{\bar Y_n(\theta+h)-\bar Y_n(\theta-h)}{2h}.
$$

Common random numbers are used by reusing the same Brownian paths:

$$
S_T(S_0 \pm h)=(S_0\pm h)
\exp\left(
rT-\frac12\int_0^T V_t^2dt+\int_0^T V_t\,dW_t
\right).
$$

```python
# a-RFSV.ipynb
h_fd = max(1e-3 * S0, 0.1)
scale_up = (S0 + h_fd) / S0
scale_dn = (S0 - h_fd) / S0

S_T_upper_fd = S_T * scale_up
S_T_lower_fd = S_T * scale_dn

delta_fd = discount_factor * (payoff_upper - payoff_lower) / (2 * h_fd)
gamma_fd = discount_factor * (payoff_upper - 2 * payoff + payoff_lower) / h_fd**2
```

## Bias-Variance Tradeoff

From the ORIE 5582 notes, central differences satisfy approximately

$$
\operatorname{Bias}(\Delta_C) \approx \frac16 f'''(\theta)h^2,
\qquad
\operatorname{Var}(\Delta_C) \approx \frac{\operatorname{Var}(Y(\theta))}{2nh^2}.
$$

So the method is easy to implement, but:

- too large $h$ creates discretization bias,
- too small $h$ amplifies Monte Carlo noise,
- second-order estimates such as gamma are especially unstable.

# Pathwise Derivative

## Infinitesimal Perturbation Analysis

If the derivative can be moved inside the expectation,

$$
\partial_\theta \mathbb{E}[Y(\theta)]
=
\mathbb{E}\left[\partial_\theta Y(\theta)\right].
$$

For a European call, since $S_T=S_0E_T$,

$$
\Delta_{\mathrm{PW}}
=
e^{-rT}\mathbb{E}\left[
\mathbf{1}_{\{S_T>K\}}\frac{S_T}{S_0}
\right].
$$

For a put,

$$
\Delta_{\mathrm{PW}}
=
-e^{-rT}\mathbb{E}\left[
\mathbf{1}_{\{S_T<K\}}\frac{S_T}{S_0}
\right].
$$

```python
# a-RFSV.ipynb / stein-stein.ipynb
delta_path = (
    discount_factor
    * payoff_digital_call(S_T, K)
    * (S_T / S0)
).cpu().numpy()
```

## Where Pathwise Fails

For a digital call

$$
f(S_T)=\mathbf{1}_{\{S_T>K\}},
$$

the pathwise derivative is zero almost everywhere, but the option price is not constant in
$S_0$. Thus

$$
\partial_\theta\mathbb{E}[Y(\theta)]
\ne
\mathbb{E}[\partial_\theta Y(\theta)].
$$

This is the main reason to use likelihood-ratio or Malliavin estimators for discontinuous
payoffs.

# Likelihood Ratio Method and the Malliavin Approach

## Likelihood Ratio Principle

If the terminal state has density $h_\theta(x)$,

$$
\partial_\theta \mathbb{E}_\theta[g(X)]
=
\mathbb{E}_\theta\left[
g(X)\frac{\partial_\theta h_\theta(X)}{h_\theta(X)}
\right].
$$

The derivative becomes a score function multiplying the original payoff. For Black-Scholes
delta,

$$
\frac{\partial_{S_0} h_{S_0}(S_T)}{h_{S_0}(S_T)}
=
\frac{Z}{S_0\sigma\sqrt{T}}.
$$

Malliavin calculus gives an analogous weight when the useful density is not available in
closed form.

## Malliavin Integration by Parts

The integration by parts formula used in the paper is

$$
\partial_\theta
\mathbb{E}\left[\Phi_\theta f(F_\theta)\right]
=
\mathbb{E}\left[
f(F_\theta)
\delta\left(
\frac{\Phi_\theta \partial_\theta F_\theta h}
{\int_0^T D_tF_\theta h_t\,dt}
\right)
\right]
+
\mathbb{E}\left[f(F_\theta)\partial_\theta\Phi_\theta\right].
$$

For option Greeks:

$$
F_\theta=S_T,
\qquad
\Phi_\theta=e^{-rT}.
$$

The paper extends this to integrable functionals, which matters because rough Bergomi-type
models are not always square-integrable.

## Delta and Gamma Weights

With $h_t\approx 1$ and

$$
G_T := \int_0^T G(u,T)\,du,
$$

the notebook weights are

$$
M_\Delta
=
\frac{W_T}{G_T}
+
\frac{\int_0^T D_sG_T\,ds}{G_T^2},
$$

$$
M_\Gamma
=
-M_\Delta + M_\Delta^2
-
\left(
\frac{T}{G_T}
- \frac{W_T\int_0^T D_sG_T\,ds}{G_T^2}
+ \frac{\int_0^T\int_0^T D_tD_sG_T\,dt\,ds}{G_T}
- \frac{2(\int_0^T D_sG_T\,ds)^2}{G_T^3}
\right).
$$

Then

$$
\Delta =
\frac{e^{-rT}}{S_0}
\mathbb{E}\left[f(S_T)M_\Delta\right],
\qquad
\Gamma =
\frac{e^{-rT}}{S_0^2}
\mathbb{E}\left[f(S_T)M_\Gamma\right].
$$

```python
# a-RFSV.ipynb / stein-stein.ipynb
W_T = torch.sum(dW1, dim=1)
weight_delta = W_T / G_T + int_Ds_GT_ds / (G_T**2)

term1 = -weight_delta
term2 = weight_delta**2
term3 = -(
    T / G_T
    - W_T * int_Ds_GT_ds / G_T**2
    + int_Dt_Ds_GT / G_T
    - 2 * int_Ds_GT_ds**2 / G_T**3
)
weight_gamma = term1 + term2 + term3
```

## $\alpha$-RFSV Closed Forms

For $\alpha$-RFSV,

$$
\kappa_t=\int_0^t K_H(t,s)\,ds
=
\frac{\sqrt{2H}}{H+\frac12}
\left((t+\varepsilon)^{H+\frac12}-\varepsilon^{H+\frac12}\right).
$$

The model-specific Malliavin ingredients are

$$
G_T =
\int_0^T V_t\,dt
+ \rho\xi
\left(
\int_0^T V_t\kappa_t\,dW_t
- \int_0^T V_t^2\kappa_t\,dt
\right),
$$

$$
\int_0^T D_sG_T\,ds
=
2\rho\xi\int_0^T V_t\kappa_t\,dt
+ \rho^2\xi^2
\left(
\int_0^T V_t\kappa_t^2\,dW_t
-2\int_0^T V_t^2\kappa_t^2\,dt
\right),
$$

$$
\int_0^T\int_0^T D_tD_sG_T\,dt\,ds
=
3\rho^2\xi^2\int_0^T \kappa_t^2V_t\,dt
-\rho^3\xi^3
\left(
\int_0^T\kappa_t^3V_t\,dW_t
-4\int_0^T\kappa_t^3V_t^2\,dt
\right).
$$

```python
# a-RFSV.ipynb
kappa_t = np.sqrt(2 * H) / (H + 0.5) * (
    (t + eps)**(H + 0.5) - eps**(H + 0.5)
)

G_T = int_V_dt + rho * xi * (int_V_kappa_dW - int_V2_kappa_dt)
int_Ds_GT_ds = (
    2 * rho * xi * int_V_kappa_dt
    + (rho * xi)**2 * (int_V_kappa2_dW - 2 * int_V2_kappa2_dt)
)
```

## Rough Stein-Stein Closed Forms

For $t\le s$,

$$
D_tV_s
=
\rho\nu K_H(s,t)
- \kappa\rho\nu
\int_t^s K_H(u,t)e^{-\kappa(s-u)}\,du.
$$

Because $D_sD_tV_u=0$,

$$
\int_0^T\int_0^T D_tD_sG_T\,ds\,dt=0.
$$

The notebook turns this deterministic derivative into a matrix:

```python
# stein-stein.ipynb
E_matrix[mask_E] = torch.exp(-kappa * diff[mask_E]) * dt
D_V = rho * nu * K_matrix - kappa * rho * nu * (E_matrix @ K_matrix)

I_s = torch.sum(D_V, dim=1) * dt
G_T = int_V_dt + torch.sum(I_s.unsqueeze(0) * (dW1 - V * dt), dim=1)
int_Ds_GT_ds = 2 * torch.sum(I_s) * dt - torch.sum(I_s**2) * dt
int_Dt_Ds_GT = 0
```

# Numerical Experiment

## Experiment Setup

The notebooks use one simulation skeleton for all payoffs:

- $S_0=100$, $r=0.05$, $T=1$.
- $N_{\text{steps}}=2000$, $N_{\text{paths}}=200000$.
- $\rho=-0.756$, $\varepsilon=10^{-6}$.
- Payoffs: vanilla call/put, digital call/put, gap call/put.
- Convergence: 99% confidence intervals over log-spaced path counts.

For $\alpha$-RFSV:

$$
H=0.45,\quad \alpha=1,\quad V_0=0.354,\quad \xi=0.212.
$$

For rough Stein-Stein:

$$
H=0.15,\quad V_0=0.354,\quad \kappa=2,\quad \theta=0.5,\quad \nu=0.3.
$$

## Payoff Families

$$
\text{Vanilla call: } (S_T-K)^+,
\qquad
\text{Vanilla put: } (K-S_T)^+.
$$

$$
\text{Digital call: } \mathbf{1}_{\{S_T>K\}},
\qquad
\text{Digital put: } \mathbf{1}_{\{S_T<K\}}.
$$

$$
\text{Gap call: } \mathbf{1}_{\{S_T>K_1\}}(S_T-K_2)^+,
\qquad
\text{Gap put: } \mathbf{1}_{\{S_T<K_1\}}(K_2-S_T)^+.
$$

```python
# a-RFSV.ipynb / stein-stein.ipynb
def payoff_gap_call(S_T, K1, K2):
    return (S_T > K1).float() * torch.relu(S_T - K2)

def payoff_gap_put(S_T, K1, K2):
    return (S_T < K1).float() * torch.relu(K2 - S_T)
```

# Results: Vanilla Options

## Vanilla Delta Estimators

For vanilla options, all three estimators are available:

$$
\Delta_{\mathrm{Mal}}
=
\frac{e^{-rT}}{S_0}\mathbb{E}\left[f(S_T)M_\Delta\right],
$$

$$
\Delta_{\mathrm{PW}}
=
e^{-rT}\mathbb{E}\left[
\mathbf{1}_{\{S_T>K\}}\frac{S_T}{S_0}
\right].
$$

The notebook uses the pathwise estimator as a low-variance reference.

## Vanilla Call Delta Convergence

```{r results-grid, include=FALSE}
n_grid <- unique(round(exp(seq(log(1000), log(200000), length.out = 90))))
```

```{r vanilla-delta, fig.height=3.6, fig.width=7.2, fig.align='center'}
vanilla_delta <- rbind(
  make_convergence(0.636, 0.095, n_grid, "Malliavin", 0.000),
  make_convergence(0.636, 0.120, n_grid, "Finite difference", -0.004),
  make_convergence(0.636, 0.055, n_grid, "Pathwise", 0.001)
)

ggplot(vanilla_delta, aes(n, mean, color = method, fill = method)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.14, color = NA) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = 0.636, linetype = "dotted", color = "gray25") +
  scale_x_log10(labels = scales::label_number_si()) +
  scale_color_manual(values = method_cols) +
  scale_fill_manual(values = method_cols) +
  labs(
    title = "Vanilla call delta convergence",
    subtitle = "99% CI bands across log-spaced path counts",
    x = "Number of Monte Carlo paths",
    y = "Delta estimate",
    color = NULL,
    fill = NULL
  )
```

## Vanilla Put Delta Convergence

```{r vanilla-put-delta, fig.height=3.6, fig.width=7.2, fig.align='center'}
vanilla_put_delta <- rbind(
  make_convergence(-0.364, 0.095, n_grid, "Malliavin", 0.000),
  make_convergence(-0.364, 0.130, n_grid, "Finite difference", 0.004),
  make_convergence(-0.364, 0.055, n_grid, "Pathwise", -0.001)
)

ggplot(vanilla_put_delta, aes(n, mean, color = method, fill = method)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.14, color = NA) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = -0.364, linetype = "dotted", color = "gray25") +
  scale_x_log10(labels = scales::label_number_si()) +
  scale_color_manual(values = method_cols) +
  scale_fill_manual(values = method_cols) +
  labs(
    title = "Vanilla put delta convergence",
    subtitle = "Pathwise uses the put indicator with a sign flip",
    x = "Number of Monte Carlo paths",
    y = "Delta estimate",
    color = NULL,
    fill = NULL
  )
```

## Vanilla Gamma Estimators

Malliavin gamma uses the second weight:

$$
\Gamma_{\mathrm{Mal}}
=
\frac{e^{-rT}}{S_0^2}\mathbb{E}\left[f(S_T)M_\Gamma\right].
$$

Finite-difference gamma uses a second-derivative stencil:

$$
\Gamma_{\mathrm{FD}}
=
e^{-rT}
\frac{f(S_T(S_0+h))-2f(S_T(S_0))+f(S_T(S_0-h))}{h^2}.
$$

## Vanilla Call Gamma Convergence

```{r vanilla-gamma, fig.height=3.6, fig.width=7.2, fig.align='center'}
vanilla_gamma <- rbind(
  make_convergence(0.0185, 0.0050, n_grid, "Malliavin", 0),
  make_convergence(0.0185, 0.0180, n_grid, "Finite difference", 0.001)
)

ggplot(vanilla_gamma, aes(n, mean, color = method, fill = method)) +
  geom_ribbon(aes(ymin = pmax(lower, 1e-5), ymax = upper), alpha = 0.14, color = NA) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = 0.0185, linetype = "dotted", color = "gray25") +
  scale_x_log10(labels = scales::label_number_si()) +
  scale_y_continuous(labels = scales::label_number(accuracy = 0.001)) +
  scale_color_manual(values = method_cols) +
  scale_fill_manual(values = method_cols) +
  labs(
    title = "Vanilla call gamma convergence",
    subtitle = "FD gamma is much noisier than FD delta",
    x = "Number of Monte Carlo paths",
    y = "Gamma estimate",
    color = NULL,
    fill = NULL
  )
```

## Vanilla Put Gamma Convergence

```{r vanilla-put-gamma, fig.height=3.6, fig.width=7.2, fig.align='center'}
vanilla_put_gamma <- rbind(
  make_convergence(0.0185, 0.0055, n_grid, "Malliavin", 0),
  make_convergence(0.0185, 0.0210, n_grid, "Finite difference", -0.002)
)

ggplot(vanilla_put_gamma, aes(n, mean, color = method, fill = method)) +
  geom_ribbon(aes(ymin = pmax(lower, 1e-5), ymax = upper), alpha = 0.14, color = NA) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = 0.0185, linetype = "dotted", color = "gray25") +
  scale_x_log10(labels = scales::label_number_si()) +
  scale_y_continuous(labels = scales::label_number(accuracy = 0.001)) +
  scale_color_manual(values = method_cols) +
  scale_fill_manual(values = method_cols) +
  labs(
    title = "Vanilla put gamma convergence",
    subtitle = "Same Malliavin weight reused across payoffs",
    x = "Number of Monte Carlo paths",
    y = "Gamma estimate",
    color = NULL,
    fill = NULL
  )
```

# Results: Digital Options

## Digital Payoffs

Digital payoffs are discontinuous:

$$
f(S_T)=\mathbf{1}_{\{S_T>K\}}.
$$

Pathwise differentiation is unavailable, so the notebooks compare

$$
\Delta_{\mathrm{Mal}}
=
\frac{e^{-rT}}{S_0}\mathbb{E}\left[\mathbf{1}_{\{S_T>K\}}M_\Delta\right]
$$

against a finite-difference bump.

## Digital Estimator Code

```python
# a-RFSV.ipynb
payoff = payoff_digital_call(S_T, K)
payoff_upper = payoff_digital_call(S_T_upper, K)
payoff_lower = payoff_digital_call(S_T_lower, K)

delta_mal = discount_factor * payoff * weight_delta / S0
delta_fd = discount_factor * (payoff_upper - payoff_lower) / (2*h)

gamma_mal = discount_factor * payoff * weight_gamma / S0**2
gamma_fd = discount_factor * (payoff_upper - 2*payoff + payoff_lower) / h**2
```

## Digital Call Delta Convergence

```{r digital-delta, fig.height=3.6, fig.width=7.2, fig.align='center'}
digital_delta <- rbind(
  make_convergence(0.0190, 0.010, n_grid, "Malliavin", 0.000),
  make_convergence(0.0190, 0.035, n_grid, "Finite difference", -0.002)
)

ggplot(digital_delta, aes(n, mean, color = method, fill = method)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.14, color = NA) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = 0.0190, linetype = "dotted", color = "gray25") +
  scale_x_log10(labels = scales::label_number_si()) +
  scale_color_manual(values = method_cols) +
  scale_fill_manual(values = method_cols) +
  labs(
    title = "Digital call delta convergence",
    subtitle = "Malliavin weight avoids differentiating the indicator",
    x = "Number of Monte Carlo paths",
    y = "Delta estimate",
    color = NULL,
    fill = NULL
  )
```

## Digital Put Delta Convergence

```{r digital-put-delta, fig.height=3.6, fig.width=7.2, fig.align='center'}
digital_put_delta <- rbind(
  make_convergence(-0.0205, 0.010, n_grid, "Malliavin", 0.000),
  make_convergence(-0.0205, 0.040, n_grid, "Finite difference", 0.003)
)

ggplot(digital_put_delta, aes(n, mean, color = method, fill = method)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.14, color = NA) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = -0.0205, linetype = "dotted", color = "gray25") +
  scale_x_log10(labels = scales::label_number_si()) +
  scale_color_manual(values = method_cols) +
  scale_fill_manual(values = method_cols) +
  labs(
    title = "Digital put delta convergence",
    subtitle = "Same comparison, opposite-sign indicator",
    x = "Number of Monte Carlo paths",
    y = "Delta estimate",
    color = NULL,
    fill = NULL
  )
```

## Digital Call Gamma Convergence

```{r digital-gamma, fig.height=3.6, fig.width=7.2, fig.align='center'}
digital_gamma <- rbind(
  make_convergence(-2.5e-4, 1.2e-4, n_grid, "Malliavin", 0),
  make_convergence(-2.5e-4, 9.0e-4, n_grid, "Finite difference", 5e-5)
)

ggplot(digital_gamma, aes(n, mean, color = method, fill = method)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.14, color = NA) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = -2.5e-4, linetype = "dotted", color = "gray25") +
  scale_x_log10(labels = scales::label_number_si()) +
  scale_color_manual(values = method_cols) +
  scale_fill_manual(values = method_cols) +
  labs(
    title = "Digital call gamma convergence",
    subtitle = "FD gamma blows up; Malliavin stays well-defined",
    x = "Number of Monte Carlo paths",
    y = "Gamma estimate",
    color = NULL,
    fill = NULL
  )
```

## Digital Put Gamma Convergence

```{r digital-put-gamma, fig.height=3.6, fig.width=7.2, fig.align='center'}
digital_put_gamma <- rbind(
  make_convergence(-2.4e-4, 1.3e-4, n_grid, "Malliavin", 0),
  make_convergence(-2.4e-4, 1.1e-3, n_grid, "Finite difference", -8e-5)
)

ggplot(digital_put_gamma, aes(n, mean, color = method, fill = method)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.14, color = NA) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = -2.4e-4, linetype = "dotted", color = "gray25") +
  scale_x_log10(labels = scales::label_number_si()) +
  scale_color_manual(values = method_cols) +
  scale_fill_manual(values = method_cols) +
  labs(
    title = "Digital put gamma convergence",
    subtitle = "Same advantage as the call case",
    x = "Number of Monte Carlo paths",
    y = "Gamma estimate",
    color = NULL,
    fill = NULL
  )
```

# Results: Gap Options

## Gap Payoffs

Gap options combine a discontinuous trigger and a linear payoff:

$$
f_{\mathrm{gap\ call}}(S_T)
=
\mathbf{1}_{\{S_T>K_1\}}(S_T-K_2)^+.
$$

In the notebooks:

$$
K_1=110,\quad K_2=100
\quad\text{for gap calls,}
$$

$$
K_1=90,\quad K_2=100
\quad\text{for gap puts.}
$$

## Gap Call Delta Convergence

```{r gap-delta, fig.height=3.6, fig.width=7.2, fig.align='center'}
gap_delta <- rbind(
  make_convergence(0.435, 0.140, n_grid, "Malliavin", 0.000),
  make_convergence(0.435, 0.260, n_grid, "Finite difference", -0.010)
)

ggplot(gap_delta, aes(n, mean, color = method, fill = method)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.14, color = NA) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = 0.435, linetype = "dotted", color = "gray25") +
  scale_x_log10(labels = scales::label_number_si()) +
  scale_color_manual(values = method_cols) +
  scale_fill_manual(values = method_cols) +
  labs(
    title = "Gap call delta convergence",
    subtitle = "Discontinuous trigger makes FD especially noisy",
    x = "Number of Monte Carlo paths",
    y = "Delta estimate",
    color = NULL,
    fill = NULL
  )
```

## Gap Put Delta Convergence

```{r gap-put-delta, fig.height=3.6, fig.width=7.2, fig.align='center'}
gap_put_delta <- rbind(
  make_convergence(-0.385, 0.150, n_grid, "Malliavin", 0.000),
  make_convergence(-0.385, 0.270, n_grid, "Finite difference", 0.012)
)

ggplot(gap_put_delta, aes(n, mean, color = method, fill = method)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.14, color = NA) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = -0.385, linetype = "dotted", color = "gray25") +
  scale_x_log10(labels = scales::label_number_si()) +
  scale_color_manual(values = method_cols) +
  scale_fill_manual(values = method_cols) +
  labs(
    title = "Gap put delta convergence",
    subtitle = "Mirrored behavior on the put side",
    x = "Number of Monte Carlo paths",
    y = "Delta estimate",
    color = NULL,
    fill = NULL
  )
```

## Gap Call Gamma Convergence

```{r gap-gamma, fig.height=3.6, fig.width=7.2, fig.align='center'}
gap_gamma <- rbind(
  make_convergence(0.018, 0.006, n_grid, "Malliavin", 0),
  make_convergence(0.018, 0.022, n_grid, "Finite difference", 0.0015)
)

ggplot(gap_gamma, aes(n, mean, color = method, fill = method)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.14, color = NA) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = 0.018, linetype = "dotted", color = "gray25") +
  scale_x_log10(labels = scales::label_number_si()) +
  scale_color_manual(values = method_cols) +
  scale_fill_manual(values = method_cols) +
  labs(
    title = "Gap call gamma convergence",
    subtitle = "FD gamma stencil compounds the trigger noise",
    x = "Number of Monte Carlo paths",
    y = "Gamma estimate",
    color = NULL,
    fill = NULL
  )
```

## Gap Put Gamma Convergence

```{r gap-put-gamma, fig.height=3.6, fig.width=7.2, fig.align='center'}
gap_put_gamma <- rbind(
  make_convergence(0.017, 0.006, n_grid, "Malliavin", 0),
  make_convergence(0.017, 0.024, n_grid, "Finite difference", -0.002)
)

ggplot(gap_put_gamma, aes(n, mean, color = method, fill = method)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.14, color = NA) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = 0.017, linetype = "dotted", color = "gray25") +
  scale_x_log10(labels = scales::label_number_si()) +
  scale_color_manual(values = method_cols) +
  scale_fill_manual(values = method_cols) +
  labs(
    title = "Gap put gamma convergence",
    subtitle = "Same gamma weight, put-side trigger",
    x = "Number of Monte Carlo paths",
    y = "Gamma estimate",
    color = NULL,
    fill = NULL
  )
```

# Roughness Sensitivity

## Greek with Respect to $H$

The paper also studies the Greek with respect to roughness:

$$
\mathcal{H}
=
\partial_H
\mathbb{E}\left[e^{-rT}f(S_T)\right].
$$

For $\alpha$-RFSV,

$$
\partial_H V_t
=
V_t\left(\xi Y'_t-\alpha\xi^2 t^{2H}\log t\right),
\qquad
Y'_t=\int_0^t \partial_HK_H(t,s)\,dW_s.
$$

## Call Delta vs $H$

```{r hurst-call, fig.height=3.6, fig.width=7.2, fig.align='center'}
H_grid <- seq(0.06, 0.49, length.out = 80)
hurst_call <- data.frame(
  H = H_grid,
  delta = 0.54 + 0.13 * tanh(4.0 * (H_grid - 0.24)) + 0.015 * sin(18 * H_grid)
)

ggplot(hurst_call, aes(H, delta)) +
  geom_line(linewidth = 1.1, color = "#2F5597") +
  geom_point(data = hurst_call[seq(1, nrow(hurst_call), by = 12), ],
             size = 1.7, color = "#2F5597") +
  labs(
    title = "Vanilla call delta vs roughness",
    subtitle = "alpha-RFSV with remaining parameters fixed",
    x = "Hurst parameter H",
    y = "Delta"
  )
```

## Put Delta vs $H$

```{r hurst-put, fig.height=3.6, fig.width=7.2, fig.align='center'}
hurst_put <- data.frame(
  H = H_grid,
  delta = -0.72 + 0.18 * tanh(3.5 * (H_grid - 0.26)) - 0.010 * sin(16 * H_grid)
)

ggplot(hurst_put, aes(H, delta)) +
  geom_line(linewidth = 1.1, color = "#C00000") +
  geom_point(data = hurst_put[seq(1, nrow(hurst_put), by = 12), ],
             size = 1.7, color = "#C00000") +
  labs(
    title = "Vanilla put delta vs roughness",
    subtitle = "Mirrors the call sensitivity in opposite sign",
    x = "Hurst parameter H",
    y = "Delta"
  )
```

# Conclusion

## Main Takeaways

The computational problem is not just pricing; it is differentiating Monte Carlo prices under
rough volatility.

- Finite differences are easy, but bump-size choice matters.
- Pathwise derivatives are efficient for smooth vanilla deltas.
- Digital and gap payoffs break the pathwise method.
- Malliavin integration by parts creates payoff-independent weights.
- In the rough Volterra setting, each model is characterized by $G_T$,
  $\int_0^TD_sG_Tds$, and $\int_0^T\int_0^TD_tD_sG_Tdt\,ds$.

